---
title: R Package Developer Essentials
description: |
  A walkthrough of the various packages, tools, resources, and axioms I have 
  encountered over my years as an R package sofware engineer. 
author:
  - name: Jimmy Briggs 
    url: https://devfolio.jimbrig.com
date: '2021-12-18'
output: 
  distill::distill_article:
    toc: true
    toc_depth: 2
    toc_float: true
bibliography: references.bib
preview: 'images/R.jpg'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(knitr.table.format = "html") 
library(tibble)
library(knitr)
library(kableExtra)
library(tufte)
```

People who program using the R language can be separated into two core
buckets:

1.  R Developers
2.  R Users

To distinguish, R developers *develop* new innovative solutions while R
Users *use* the tools and packages that the developers created. This
article is aimed towards the development side of the spectrum and
attempts to provide a comprehensive, curated toolbox for the R
Developer, specifically in regard to developing R Packages.

## Pre-Requisite Readings

To start, R package developers need to be familiar with the fundamentals
of package development, and therefore should be familiar with some of
the highest regarded resources available to read.

Anyone who is serious about developing production grade R packages needs
to be familiar with most of, if not all of the following resources and
guides:

```{r}
data <- tibble::tribble(
  ~Resource,                                                                ~Source,
  "[Writing R Extensions](http://cran.fhcrc.org/doc/manuals/R-exts.html)", "[CRAN Manuals](https://cran.r-project.org/manuals.html)",
  "[R Packages](http://r-pkgs.had.co.nz/)",                                    "[Hadley Wickham](http://hadley.nz/)",
  "[R Package Primer](http://kbroman.org/pkg_primer/)",               "[Karl Broman](https://kbroman.org/)",
  "[Package Guidelines](http://www.bioconductor.org/developers/package-guidelines/)",             "Bioconductor",
  "[rOpenSci Packages Developer Guide](https://devguide.ropensci.org/index.html)", "[rOpenSci](https://ropensci.org/)"
) 

data %>% 
  knitr::kable(align = "c", booktabs = TRUE, 
               caption = "R Package Developer Essential Resources") %>%
  kableExtra::kable_material(c("striped"), full_width = TRUE) %>%
  kableExtra::row_spec(0, bold = TRUE, color = "white", background = "black")
```

<details><summary>View Citations</summary><p>

[@wickhamb; @whypacka; @vidoni; @initiativea; @anintro; @gandrud2015;
@glennie2020; @owen-the; @riederera; @spector2004; @teama; @whypackb;
@wickhamc; @zhua]

</p></details></p></summary></details>

### Writing R Extensions Manual

The [Writing R Extensions
Manual](http://cran.fhcrc.org/doc/manuals/R-exts.html) is perhaps the
most crucial resource listed above, and has even been considered the
*Bible of R Package Development*.

However, it is very exhaustive and not the most visually-appealing
write-up.

As Hadley puts it in his book [Writing R Packages](): [@wickham]

> "The best resource for the official details of package development is
> always the official writing R extensions manual. However, this manual
> can be hard to understand if you're not already familiar with the
> basics of packages. It's also exhaustive, covering every possible
> package component, rather than focusing on the most common and useful
> components, as this book does. Writing R extensions is a useful
> resource once you've mastered the basics and want to learn what's
> going on under the hood."
>
> `r tufte::quote_footer('--- Hadley Wickham')`

Thanks to [Colin Fay](https://colinfay.me/), a more elegant version of
the original manual has been created as a [bookdown]() site and
published online at <https://colinfay.me/writing-r-extensions>.

This resource is highly encouraged for anyone taking R Package
Development seriously.

*Note: The other manuals listed on the [CRAN
Manuals](https://cran.r-project.org/manuals.html) website contain a lot
of hidden gems that are often overlooked by R developers. These
resources contain some of the most crucial, foundational knowledge that
anyone using R should eventually be made aware of, therefore I highly
recommend you check those out in addition to Writing R Extensions*.

## Package Development Fundamentals

R Package development can be broken down into the following fundamental
areas of development:

``` mermaid
```
